%!TEX root = ../main.tex
\chapter{Tate's Linear Algebra}\label{ch:tate-linear-algebra}
In this chapter we explore linear topologies on vector spaces in order to introduce Tate spaces and their structure. Tate spaces will be central in the definition of abstract residues in Chapter 2 and the study of algebraic curves in Chapter 3. We follow definitions in \cite{Chiral} closely but not religiously.
\section{Linear topologies}
Let $k$ be a field. From now on, a vector space will always mean a $k$-vector space.
\begin{definition}\label{def:linear_topology}
A \textbf{linear topology} on a vector space $V$ is a separated (Hausdorff) topology which is invariant under translations and which admits a base of open neighborhoods of zero consisting of vector subspaces. A vector space equipped with a linear topology will be referred as \textbf{linearly topologized}.
\end{definition}
If we endow $k$ with the discrete topology then $V$ will become a topological vector space. From now on, endow $k$ with the discrete topology. \\
Linear topologies behave nicely under basic topological operations.
\begin{theorem}\label{thm:linear_topologies_properties}
Let $V$ be a linearly topologized vector space. Then
	\begin{enumerate}[label = (\alph*)]
		\item If $W \subseteq V$ is a vector subspace then $W$ is linearly topologized as well.
		\item If $W \subseteq V$ is a closed vector subspace then $V/W$ is linearly topologized under its quotient topology.
		\item If $\{V_{\alpha}\}_{\alpha}$ is a collection of linearly topologized vector spaces its product $\prod_{\alpha} V_{\alpha}$ (in its product topology) and its direct sum $\bigoplus_{\alpha} V_{\alpha}$ (as a subspace of the product) are linearly topologized.
		\item If $W$ is a vector subspace of $V$, then its topological closure $\overline{W}$ also is a vector subspace of $V$.
	\end{enumerate}
\end{theorem}
\begin{proof}
	If $\mathscr{U}$ is a system of neighborhoods around zero consisting of vector subspaces in $V$ then $\{U\cap W\mid U \in \mathscr{U}\}$ is a system of neighborhoods around zero consisting of vector subspaces in $W$. For (b), let $\pi\colon V \to V/W$ be the quotient map. Since $\pi$ is open and surjective the image of a local base is a local base; moreover, the image of a vector subspace under $\pi$ is a vector subspace. In addition, since $W$ is closed it follows that $V/W$ is Hausdorff.  Now, for (c) let $\{U_{\alpha, \beta}\}_{\beta}$ be a local base of zero in $V_{\alpha}$ of vector subspaces, the products $U_{\alpha_{1}, \beta_{1}} \times \ldots \times U_{\alpha_{n}, \beta_{n}} \times \prod_{\gamma} V_{\gamma}$, where $\gamma$ ranges over $\alpha \neq \alpha_{1}, \ldots, \alpha_{n}$, for any set $\{(\alpha_{1}, \beta_{1}, \ldots, \alpha_{n}, \beta_{n})\}$ form a fundamental system of neighborhoods around zero in $\prod_{\alpha} V_{\alpha}$ of open vector subspaces. Note that since $\bigoplus_{\alpha} V_{\alpha} \subseteq \prod_{\alpha} V_{\alpha}$ is a vector subspace (c) follows from (a). Finally, for (d), suppose $x,y\in \overline{W}$, then, for every open vector subspace $U$, $(x + U)\cap W \neq \emptyset$ and $(y + U)\cap W \neq \emptyset$, therefore for every $\alpha, \beta \in k$ we have $(\alpha x + U)\cap W \neq \emptyset$ and $(\beta y + U)\cap W \neq \emptyset$. Hence, $(\alpha x + \beta y + U)\cap W\neq \emptyset$ for every open vector subspace $U$ and every pair $\alpha, \beta\in k$. It follows (d).
\end{proof}
\begin{corollary}\label{cor:limits-and-colimits-in-lintop-category}
	Let $\mathsf{LinTop}_{k}$ denote the category of linearly topologized vector spaces over $k$, where morphisms are given by continuous linear homomorphisms. Then, limits exist in $\mathsf{LinTop}_{k}$. 
\end{corollary}
\begin{proof}
	By \cref{thm:linear_topologies_properties} it follows that kernels and arbitrary products exist in $\mathsf{LinTop}_{k}$; therefore, limits exist in $\mathsf{LinTop}_{k}$. 
\end{proof}

\begin{proposition}\label{prop:finite_dimensional_linear_topologies}
A finite dimensional linearly topologized vector space $V$ is discrete.
\end{proposition}
\begin{proof}
	Since $V$ is Hausdorff, it follows that
	\[
		\bigcap_{U \in \mathscr{U}}U = \{0\},
	\]
	for $\mathscr{U}$ a system of neighborhoods of zero consisting of vector subspaces of $V$. Since $V$ is finite dimensional there exist $U_{1}, \ldots, U_{n} \in \mathscr{U}$ such that
	\[
		\bigcap_{U \in \mathscr{U}}U = U_{1}\cap \ldots \cap U_{n} = \{0\}.
	\]
	 Therefore, $\{0\}$ is open. This implies that $V$ is separated.
\end{proof}
\begin{proposition}\label{prop:characterization-of-closure}
	Let $V$ be a linearly topologized vector space and $W$ a vector subspace of $V$. Then
	\[
		\overline{W} = \bigcap_{U\in \mathscr{U}} (W + U).
	\]
	for a system of neighborhoods of zero $\mathscr{U}$ consisting of vector subspaces of $V$.
\end{proposition}
\begin{proof}
	Since every $W + U$ is open, it is also closed. Hence the intersection is closed. If $x\in V - \overline{W}$, then $(x + U)\cap W = \{0\}$ for some $U \in \mathscr{U}$. Hence, $x\notin W + U$. This proves the proposition.
\end{proof}
\subsection{Commensurability}
We introduce a partial order on the set of vector subspaces of a vector space $V$.
\begin{definition}\label{def:commensurability}
	For vector subspaces $A$ and $B$ of a vector space $V$ we say that $A \prec B$ if the quotient $A/(A\cap B) \cong (A+B)/B$ is finite dimensional (or equivalently $A \subseteq B + W$ for some finite dimensional $W$). In addition, we say that $A$ and $B$ are \textbf{commensurable} (denoted $A \sim B$) if $A \prec B$ and $B \prec A$.
\end{definition}
Observe that $A \sim B$ if and only if $(A+B)/(A\cap B) \cong A/(A\cap B) \oplus B/(A \cap B)$ is finite dimensional. We will constantly refer to a vector space $V$ being finite dimensional as $V \sim 0$.
\begin{proposition}\label{prop:equivalence-relation}
	Let $V$ be a vector spaces and $A,B$ and $C$ be vector subspaces, then:
	\begin{enumerate}[label = (\alph*)]
		\item If $A \sim B$ and $B \sim C$ then
		\[
			\frac{A+B+C}{A \cap B \cap C} \sim 0
		\]
		\item If $A \prec B$ and $B \prec C$ then $A \prec C$. Moreover, commensurability is an equivalence relation.
	\end{enumerate}
\end{proposition}
\begin{proof}
	Consider the following exact sequences
	\[
		0 \to \frac{A\cap B}{A \cap B \cap C} \to \frac{B}{B \cap C}, 
	\]
	and,
	\[
		0 \to \frac{A\cap B}{A \cap B \cap C} \to \frac{A+B}{A \cap B \cap C}
		\to \frac{A+B}{A \cap B} \to 0
	\]
	induced by inclusions. The first inclusion plus the fact that $B \sim C$ imply that $(A\cap B)/(A \cap B \cap C)$ is finite dimensional. Now, since $A \sim B$ it follows that $(A+B)/(A \cap B)$ is finite dimensional. Therefore, using the second exact sequence we conclude that $(A+B)/(A \cap B \cap C)$ is finite dimensional. A symmetrical argument shows that $(B+C)/(A \cap B \cap C) \sim 0$. These prove (a). For (b), the inclusion
	\[
		0 \to \frac{A+C}{A\cap C} \to \frac{A+B+C}{A \cap B \cap C}
	\]
	plus (a) implies transitivity. 
\end{proof}
Now, we state and prove some useful properties on the relation $\prec$.
\begin{lemma}\label{lemm:properties-order-well-behaved-under-operations}
Let $V$ be a vector space and $A,B$ vector subspaces of $V$. Then
\begin{enumerate}[label = (\alph*)]
	\item If $A \subseteq B$ it follows $A \prec B$.
	\item If $A \prec B$ then $f(A) \prec f(B)$ for any $k$-linear map $f$
	\item Let $\{A_{i}\}_{i=1}^{m}$ and $\{B_{j}\}_{j=1}^{n}$ be two collections of vector subspaces of $V$. Then,
	\[
		\sum_{i=1}^{m} A_{i} \prec \bigcap_{j=1}^{n} B_{j} \iff A_{i} \prec B_{j}\text{  for all } i \text{ and } j.
	\]
\end{enumerate}
\end{lemma}
\begin{proof}
	First, (a) is immediate from the definition of $\prec$. Second, for (b) the map $f$ factors as
	\[
		A/(A\cap B) \to f(A)/(f(A)\cap f(B)) \to 0
	\]
	Finally, for (c), if $\sum_{i=1}^{m} A_{i} \prec \bigcap_{j=1}^{n} B_{j}$ holds then by (a) above, for all $i$ and $j$ we have
	\[
		A_{i} \prec \sum_{i=1}^{m} A_{i} \prec \bigcap_{j=1}^{n} B_{j} \prec B_{j}
	\]
	On the other hand, if $A_{i} \prec B_{j}$ for all $i$ and $j$ then there exist finite dimensional subspaces $W_{ij}$ such that $A_{i} \subseteq B_{j} + W_{ij}$ for all $i$ and $j$. Therefore,
	\[
		\sum_{i=1}^{m} A_{i} \subseteq \bigcap_{j=1}^{n} B_{j} + \sum_{i=1}^{m} \sum_{j=1}^{n} W_{ij}.
	\]

\end{proof}
Next, we consider another useful lemma.
\begin{lemma}\label{lemm:commensurability-addition-and-intersection}
	Let $A,B,A',B'$ be vector subspaces of a vector space $V$ and suppose that $A \sim A'$ and $B \sim B'$. Then $A + B \sim A' + B'$ and $A \cap B \sim A' \cap B'$.
\end{lemma}
\begin{proof}
	The following exact sequence 
	\small
	\[
		0 \to \frac{A + A' + B + B'}{A\cap A'\cap B\cap B'} \to \frac{A + A'}{A \cap A'} \oplus \frac{B + B'}{B \cap B'} \to \frac{A + A' + B + B'}{(A\cap A') + (B\cap B')} \to 0
	\]
	\normalsize
	plus $A \sim A'$ and $B \sim B'$ imply that both spaces
	\[
		\frac{A + A' + B + B'}{A\cap A'\cap B\cap B'} \quad\text{and,}\quad
		 \frac{A + A' + B + B'}{(A\cap A') + (B\cap B')}
	\]	
	are finite dimensional. Since, $(A + A' + B + B')/(A+A')\cap(B+B')$ is a quotient of the second space and $((A \cap A') + (B \cap B'))/((A\cap A')\cap(B\cap B'))$ is a subspace of the first space we can conclude $A + B \sim A' + B'$ and $A \cap B \sim A' \cap B'$.
\end{proof}
If we consider the set of equivalence classes of $\sim$ on a vector space $V$ then $\prec$ is a partial order on it and by \cref{lemm:commensurability-addition-and-intersection} above it inherits operations $\cap$ and $+$.
\subsection{Linear compactness}
\begin{definition}\label{def:linear_compactness}
	Let $V$ be a linearly topologized vector space. A closed vector subspace $L \subseteq V$ is \textbf{linearly compact} (respectively \textbf{linearly cocompact}) if for every open vector subspace $U$ we have $L \prec U$ (respectively $V/(L+U) \sim 0$). 
\end{definition}
\begin{remark}\label{rem:lefschetz-linear-compactness}
	Linear compactness was introduced by S. Lefschetz in his influential \cite{Lefschetz-Alg-Top} using different terms. He defined a coset of $V$ (also called linear variety) to be a set $x + W$ where $x\in V$ and $W$ is a subspace $V$. Then, he defined a linearly topologized vector space $V$ to be linearly compact if for every collection of closed cosets $X_{\alpha}$ having the finite intersection property follows that $\bigcap_{\alpha} X_{\alpha} \neq \emptyset$. In these terms, linear compactness seems like a natural generalization of compactness for linearly topologized vector spaces. We extend this discussion in \cref{rem:lef-vs-beilinson}.
\end{remark}
Linear compactness behaves just as compactness if one uses the correct words.
\begin{theorem}\label{thm:linear_compactness_properties}
	Let $V$ be a linearly compact vector space, then
	\begin{enumerate}[label = (\alph*)]
		\item If $A \subseteq V$ is a vector subspace satisfying $A \prec U$ for all open vector subspaces $U$ of $V$ then $\overline{A}$ is linearly compact.
		\item If $f\colon V \to W$ is a continuous linear homomorphism then $\overline{f(V)}$ is linearly compact.
		\item If $V$ is discrete then $V \sim 0$.
		\item Every closed vector subspace of $V$ is linearly compact.
		\item (Tychonov) If $\{V_{\alpha}\}_{\alpha}$ is a collection of linearly compact vector spaces then its product $\prod_{\alpha} V_{\alpha}$ and its direct sum $\bigoplus_{\alpha}V_{\alpha}$ are linearly compact.
	\end{enumerate}
\end{theorem}
\begin{proof}
	Let $U$ be any open vector subspace of $V$. Then, $A + U$ is closed, that is $A + U = \overline{A + U} \supseteq \overline{A} + U \supseteq A + U$, thus, $\overline{A} + U = A + U$. Since, $(A + U)/U \sim 0$ it follows that $(\overline{A} + U)/U \sim 0$. 

	For (b), since $f$ is a continuous linear map $V \prec f^{-1}(U)$ for all $U$ open vector subspaces of $W$. Hence by \cref{lemm:properties-order-well-behaved-under-operations} $f(V) \prec U$ for all open vector subspaces $U$ of $W$. The previous observation and (b) yield (a). If $V$ is discrete, then $\{0\}$ is an open vector subspace of $V$; therefore, $V$ is finite dimensional.

	For (d), if $A \subseteq V$ is a closed vector subspace, and $V \prec U$ for all open vector subspaces $U$ by \cref{lemm:properties-order-well-behaved-under-operations} we get $A \prec U$. 

	Finally, for (d), it is enough proving for open vector subspaces $U = \prod_{\beta} U_{\beta} \times \prod_{\gamma} V_{\gamma}$ where $\beta$ ranges over a finite set, $\gamma$ ranges over $\alpha \neq \beta$ and $U_{\beta}$ is an open vector subspace of $V_{\beta}$. Then, the quotient
	\[
		\prod_{\alpha}V_{\alpha} / U \cong \prod_{\beta} V_{\beta}/U_{\beta}
	\]
	where $\cong$ is a topological and algebraic isomorphism. Since $V_{\alpha}$ is linearly compact for all $\alpha$ and $\beta$ ranges over a finite set we conclude that $\prod_{\alpha} V_{\alpha} / U$ is finite dimensional; therefore, $\prod_{\alpha} V_{\alpha}$ is linearly compact. The proof is analogous for $\bigoplus_{\alpha} V_{\alpha}$.
\end{proof} 
\subsection{Completion}
In this subsection we expose with little detail the properties of completion. A general reference for completion of arbitrary modules over a commutative ring is \cite{Matsumura} Section 8. 
%\begin{definition}\label{def:net-in-a-topological-vector-space}
%mathscr{U}$
%\[
%x_{U'} - x \in U.
%\]
%\end{definition}
%\begin{remark}\label{rem:nets}
%$\mathcal{D} = \mathscr{U}^{\text{op}}$.
%\end{remark}
%We now introduce Cauchy nets and completion.
%\begin{definition}\label{def:cauchy-nets}
%\subseteq U$ we have $x_{U_{1}} - x_{U_{2}} \in U$. 
%\end{definition}
%
%\begin{definition}\label{def:completion}
%textbf{complete} if 
%\[
%V \cong \varprojlim_{U \in \mathscr{U}} V/U
%\]
%identifying the inverse limit as a subspace of the product.
%\end{definition}
%\textcolor{red}{Complete this! (no pun intended) }
\begin{definition}\label{def:completion}
	Let $V$ a linearly topologized vector space. Let $\mathscr{U} = \{U_{\lambda}\}_{\lambda\in \Lambda}$ be a system of neighborhoods of zero consisting of open vector subspaces of $V$ indexed by $\Lambda$, so that $\lambda < \mu \iff U_{\lambda} \supseteq U_{\mu}$. Then, the inverse limit
	\[
		\widehat{V} = \varprojlim_{U\in \mathscr{U}} V/U = \varprojlim_{\lambda\in \Lambda} V/U_{\lambda}
	\]
	is the \textbf{completion} of $V$. Observe that existence of such limit is guaranteed by \cref{cor:limits-and-colimits-in-lintop-category}. Recall that $\widehat{V}$ can be described as a subspace of the product 
	\[
		\widehat{V} = \{(\overline{x_{\lambda}})_{\lambda}\in \prod_{\lambda\in \Lambda} V/U_{\lambda}\colon  \overline{x_{\mu}} = \pi_{\mu}^{\lambda}(\overline{x_{\lambda}}) \text{ for all } \mu \leq \lambda \},
	\] 
	where $\pi_{\mu}^{\lambda}\colon V/U_{\lambda} \to V/U_{\mu}$ is the induced map by the projection $\pi_{\mu}\colon V \to V/U_{\mu}$. In this way $\widehat{V}$ carries a natural topology as a subspace of the product $\prod_{U\in \mathscr{U}} V/U$.  There is always a natural map
	\[
		V \to \widehat{V}
	\]
	induced by the quotient maps $\pi_{\mu}$. This map is continuous and its image is dense in $\widehat{V}$. If the map $V \to \widehat{V}$ is a topological isomorphism we say that $V$ is \textbf{complete}. Let $\mathscr{U} = \{U_{\lambda}\}_{\lambda\in \Lambda}$ and $\mathscr{U}' = \{U'_{\lambda'}\}_{\lambda' \in \Lambda}$ be two different systems of neighborhoods of consisting of vector subspaces in $V$. Then for all $\lambda \in \Lambda$ there exists $\lambda'\in \Lambda'$ such that $U_{\lambda'} \subseteq U_{\lambda}$ and for all $\mu'\in \Lambda'$ there exists $\mu\in \Lambda$ such that $U_{\mu}\subseteq U_{\mu'}$. This implies that $\varprojlim_{U\in \mathscr{U}} V/U$ and $\varprojlim_{U'\in \mathscr{U}'} V/U'$ satisfy the same universal property. Therefore, they are canonically isomorphic. In other words, the completion does not depend on the choice of filtration $\mathscr{U}$.
\end{definition}
%\begin{proposition}\label{prop:completeness-is-exact-functor}
%Let $V$ be a linearly topologized vector space and $W$ a vector subspace. Then
%\begin{enumerate}[label = (\alph*)]
%\item The sequence
%\[
%0 \to \widehat{W} \to \widehat{V} \to \widehat{V/W} 
%\]
%is exact. 
%of vector subspaces of $V$, then the sequence
%\[
%0 \to \widehat{W} \to \widehat{V} \to \widehat{V/W} \to 0
%\]
%is exact. In particular, $\widehat{V}/\widehat{W} = \widehat{V/W}$.
%\end{enumerate}
%\end{proposition}
%\begin{proof}
%Observe that exactness of (a) follows from exactness of the sequences
%\[
%0 \to W/(W \cap U) \to V/U \to V/(W + U) \to 0
%\]
%Therefore, the sequence 
%\[
%0 \to W/(W \cap U_{n}) \to V/U_{n} \to V/(W + U_{n}) \to 0
%\]
%induces an exact sequence 
%\[
%0 \to \widehat{W} \to \widehat{V} \to \widehat{V/W} \to 0.
%\]
%For detalis, see Theorem 8.1 in \cite{Matsumura}.
%\end{proof}

\section{Tate spaces}
\subsection{Lattices}
\begin{definition}\label{def:c-lattice}
	If $V$ is a linearly topologized vector space we say that a \textbf{c-lattice} is an open linearly compact subspace of $V$, \textit{dually} a discrete linearly cocompact subspace is a \textbf{d-lattice}.
\end{definition}
First, we prove that existence of a c-lattice in a linearly topologized vector space is equivalent to existence of a d-lattice.
\begin{proposition}\label{prop:c-lattice-iff-d-lattice}
	A linearly topologized vector space $V$ contains a c-lattice if and only if it contains a d-lattice. 
\end{proposition}
\begin{proof}
	Suppose $L$ is a c-lattice in $V$. Choose any direct complement $D$ of $L$, that is, $V = L \oplus D$. Since $L$ is open, then $D$ is discrete as $D\cap L = 0$, thus ${0}$ is open in $D$. Moreover, $D$ is closed it is the kernel the projection $V \to L$ (which is continuous because $L$ is open). Finally, we check that $D$ is linearly cocompact: let $U$ be any open vector subspace of $V$, the composition $L \hookrightarrow V \twoheadrightarrow V/(D+U)$ induces a surjection 
	\[
		L/(L \cap U) \twoheadrightarrow V/(D+U).
	\]
	Since $\dim L / (L \cap U) < \infty$ we conclude that $\dim V/(D+U) < \infty$. \\
	Now, suppose $D$ is a d-lattice. Thus, there exists an open vector subspace $U$ such that $U \cap D = 0$. This time, choose $L$ a direct complement for $D$ containing $U$. Then, the projection $V \to D$ is continuous because $U$ is mapped to zero. Therefore, $L$ is open. Now, we prove that $L$ is linearly compact. Let $U$ be any open vector subspace. Then, the composition $V \twoheadrightarrow L \twoheadrightarrow L/(L \cap U)$ induces a surjection
	 \[
	 	V/(D + (L \cap U)) \twoheadrightarrow L/(L \cap U).
	 \]
	 Since both $L$ and $U$ are open, so is $L\cap U$. Therefore, $\dim V/(D + (L \cap U)) < \infty$. It follows that $\dim L/(L \cap U) < \infty$ and $L$ is linearly compact.  
\end{proof}
\begin{remark}\label{up-to-finite-dimension}
	Note that in the proof of \cref{prop:c-lattice-iff-d-lattice} it is not strictly necessary to choose a direct complement, one can choose a direct complement up to finite dimension; that is, $L + D \sim V$ and $L \cap D \sim 0$. 
\end{remark}
%We now give a characterization of lattices in terms of the relation $\prec$.
%\begin{proposition}\label{prop:maximality-lattices}
	%A linearly compact subspace is a c-lattice if and only if it is maximal among the set of linearly compact sets ordered by $\prec$. 
%\end{proposition}
%\begin{proof}
	%\textcolor{red}{this one needs some thinking}
%\end{proof}
%not necessary
\begin{proposition}\label{prop:lattices-are-basis}
	If $V$ admits a c-lattice, then the set of c-lattices constitutes a system of neighborhoods of zero consisting of mutually commensurable vector subspaces.
\end{proposition}
\begin{proof}
	If $L$ and $L'$ are two c-lattices in $V$ then $L \prec L'$ and $L' \prec L$ because both are open. Therefore, all c-lattices are commensurable. Moreover, if $U$ is any open vector subspace and $L$ is a c-lattice we claim that $L \cap U$ is a c-lattice as well. Indeed, let $U'$ be any open vector subspace, then $L\cap U \prec L\prec U'$. In addition, since $L$ and $U$ are open, so is $L \cap U$. Hence, $L \cap U \subseteq U$ is a c-lattice. This proves the statement.
\end{proof}
We are now ready to introduce the definition of a Tate space.
\begin{definition}\label{def:tate-vector-space}
	A linearly topologized vector space $V$ is a \textbf{Tate space} if it is complete and admits a c-lattice. By \cref{prop:lattices-are-basis} and the observation in \cref{def:completion} it follows that
	\[
		V \cong \varprojlim_{L \in \mathscr{U}} V/L
	\]
	where $\mathscr{U}$ runs through all c-lattices of $V$.
\end{definition}
\begin{example}\label{ex:tate-spaces}
We give some examples of Tate spaces.
\begin{enumerate}[label = (\alph*)]
	\item Any vector space endowed with the discrete topology is a Tate space.
	\item If $\{V_{\alpha}\}_{\alpha}$ is any projective system of finite dimensional vector spaces (thus, each one endowed with the discrete topology by \cref{prop:finite_dimensional_linear_topologies}), let $V$ be their inverse limit endowed with the inverse limit topology. We claim that this is a linearly compact space. Indeed, if we realize $V$ as a subspace of the product $\prod_{\alpha} V_{\alpha}$, a basic open vector subspace in $V$ is given by
	\[
	\{(x_{\alpha})_{\alpha} \in V \colon x_{\alpha_{1}} = x_{\alpha_{2}} = \cdots = x_{\alpha_{n}} = 0\}
	\] for some finite collection of indices $\alpha_{1}, \ldots, \alpha_{n}$. Hence, the quotient of $V$ by any basic open vector subspace is a vector subspace of a finite product of $V_{\alpha}$. Since all $V_{\alpha}$ are finite dimensional, we conclude that $V$ is linearly compact. Observe that a system of neighborhoods consisting of vector subspaces of $V$ is the collection of kernels of the projections $V \to V_{\alpha}$. Therefore, $V$ is complete.

	\item Let $V = k\left((t)\right)$ with the topology generated by letting $t^{n}k\left[[t]\right]$ for $n \in \Z$ be a system of neighborhoods of zero. Then, $V = k\left[[t]\right] \oplus t k[t^{-1}]$ where $k\left[[t]\right]$ is the completion of $k[x]$ in the $\left\langle x\right\rangle$-adic topology, hence by the previous item linearly compact and, since it is open is a c-lattice. By the argument given in \cref{prop:c-lattice-iff-d-lattice} $t k[t^{-1}]$ is a d-lattice. Therefore, $V$ is a Tate space that is neither linearly compact nor discrete.
	
	\end{enumerate}

\end{example}
A closer look at the previous examples motivates the following proposition:
\begin{proposition}\label{prop:characterization-of-tate-spaces-as-inverse-limits}
	A linearly topologized vector space $V$ is a Tate space if and only if it there exists a collection $\mathscr{U}$ in $V$ of mutually commensurable open vector subspaces in $V$ such that the natural map
	\[
		V \to \varprojlim_{U \in \mathscr{U}} V/U
	\]
	is a topological isomorphism. In particular, every Tate space arises in the following way: Let $V$ be a $k$-vector space endowed with a collection $\mathscr{U}$ of vector subspaces satisfying the following conditions:
	\begin{enumerate}[label = (\roman*)]
		\item $\mathscr{U}$ filters down to $0$ and up to $V$.
		\item Every two subspaces in $\mathscr{U}$ are mutually commensurable.
		\item The natural map
		\[
			V \to \varprojlim_{U\in \mathscr{U}} V/U
		\]
		is an isomorphism. Then, $(V, \mathscr{U})$ becomes a Tate space by imposing (iii) to be a topological isomorphism letting the quotient $V/U$ be discrete for all $U \in \mathscr{U}$.
	\end{enumerate}
	
\end{proposition}
\begin{proof}
	By the observation in \cref{def:tate-vector-space} such collection is simply the collection of c-lattices in $V$. Now, suppose that such collection exists in $V$. Then, $\mathscr{U}$ is a system of neighborhoods of zero consisting of mutually commensurable open vector subspaces in V, any $U \in \mathscr{U}$ is a c-lattice and $V$ is complete.
\end{proof}


\subsection{Duality}
If $V$ is a Tate space we consider the following topology on the dual space $V^{*}$ (where by dual space we mean topological dual). Open vector subspaces are given by
\[
	L^{\perp} = \{\phi\in E^{*} \colon \phi\lvert_{L} = 0\}
\]
where $L$ is a linearly compact subspace. Equivalently, one can define open vector subspaces in $E^{*}$ to be $D^{*}$ where $D$ a direct complement of a linearly compact vector subspace $L$ in $E$ (in this case $D^{*} \hookrightarrow E^{*}$ using the decomposition $L\oplus D$).  \\
First, we prove that the word \emph{dually} in \cref{def:linear_compactness} actually makes sense. 

\begin{lemma}\label{lemm:duality-d-lattice-c-lattice}
	Duality interchanges linearly compact with discrete spaces and vice-versa. 
\end{lemma}
\begin{proof}
	If $L$ is a linearly compact vector space, then $L^{\perp}$ is open in $L^{*}$, thus $L^{*}$ is discrete. If $D$ is discrete, then $D \cong k^{\oplus \Lambda}$ for some $\Lambda$ and endowing $k^{\oplus \Lambda}$ with the discrete topology the previous isomorphism is a homeomorphism as well. Moreover, since $D$ is discrete every linear functional is continuous. Using \cref{cor:limits-and-colimits-in-lintop-category} and the well known identity (where maps are isomorphisms in $\mathsf{LinTop}_{k}$)
	\[
		(k^{\oplus \Lambda})^{*} = \Hom_{k}(k^{\oplus \Lambda}, k) \cong \prod_{\Lambda} \Hom_{k}(k, k) \cong \prod_{\Lambda} k
	\]
	we get the desired result by Tychonov's theorem in \cref{thm:linear_compactness_properties}. 
\end{proof}
\begin{remark}\label{rem:dual-of-discrete-is-complete}
	A closer look at the proof of the previous lemma indicates that the dual space of a discrete space is always complete.
\end{remark}
\begin{proposition}\label{prop:dual-space-is-tate}
	If $V$ is a Tate space then $V^{*}$ is also a Tate space.
\end{proposition}
\begin{proof}
	If we decompose $V = L \oplus D$ where $L$ is a c-lattice and $D$ a d-lattice then $V^{*} \cong L^{*} \oplus D^{*}$ and by \cref{lemm:duality-d-lattice-c-lattice} $L^{*}$ is discrete and $D^{*}$ is linearly compact. Observe that $D^{*}$ is open in $V^{*}$ since it is the kernel of the projection $V^{*} \to V^{*}/L^{\perp}$ and $V^{*}/L^{\perp}$ is discrete by the description of our topology in the dual $V^{*}$. Since $L^{*}$ is discrete, then it is complete. Moreover, by \cref{rem:dual-of-discrete-is-complete} $D^{*}$ is complete. Hence, so is $V^{*}$.
\end{proof}
We are now ready to prove the duality theorem for Tate spaces.
\begin{theorem}\label{thm:self-duality}
	For a Tate space $V$ the canonical map $V \to V^{**}$ is an isomorphism.
\end{theorem}
\begin{proof}
	It is enough to prove it for complete linearly compact spaces and discrete spaces, as every Tate space can be decomposed into a direct sum of a c-lattice and a d-lattice. First, suppose $D$ is a discrete vector space. Then, the canonical map
	\[
		\operatorname{ev}\colon D \to D^{**}
	\]
	is open and continuous because $D$ and $D^{**}$ are both discrete by \linebreak \cref{lemm:duality-d-lattice-c-lattice}. Moreover, it is injective, because for every nonzero $x \in D$ there exists a linear continuous functional $\phi\in D^{*}$ such that $\phi(x)\neq 0$. Finally, we prove surjectivity. Let $\psi \in D^{**}$. Since $\ker \psi$ is open it contains a basic open vector subspace $A^{\perp}$ such that $A \subseteq D$ is a linearly compact subspace. Since $D^{*}$ is linearly compact it follows that $D^{*} \sim A^{\perp}$, that is, the quotient $D^{*}/A^{\perp}$ is finite dimensional. Recall that the inclusion $\iota\colon A \to D$ induces an isomorphism $D^{*}/A^{\perp} \to A^{*}$ which is a homeomorphism since both spaces are discrete. We can factor $\psi$ so that the following diagram commutes
	\[
	\begin{tikzcd}
		D^{*} \arrow[r, "\psi"] \arrow[d] & k \\
		D^{*}/A^{\perp} \arrow[ru, dashed, "\tilde{\psi}"] \arrow[d, swap, "\cong"] & \\
		A^{*} \arrow[ruu, dashed, swap, bend right, "\overline{\psi}"]
	\end{tikzcd}
	\]
	However, $A^{*}$ is finite dimensional, therefore, there exists $a \in A$ such that $\overline{\psi} = \operatorname{ev}_{a}$ as maps $A^{*}\to k$. Moreover, since $A^{\perp} \subseteq \ker\psi$ we conclude that $\psi = \operatorname{ev}_{a}$ as maps $D^{*} \to k$. This implies surjectivity. Thus $D \to D^{**}$ is an isomorphism of topological vector spaces. \\

	Now, suppose $L$ is a complete linearly compact space. We check first that the map 
	\[
		\operatorname{ev}\colon L \to L^{**}
	\]
	is continuous. Let $A^{\perp}$ be an open vector subspace in $L^{**}$ where $A \subseteq L^{*}$ is a linearly compact subspace. By \cref{lemm:duality-d-lattice-c-lattice} $L^{*}$ is discrete, hence $A$ is finite dimensional. Suppose that $A = \operatorname{span}(\phi_{1},\ldots,\phi_{n})$ for some $\phi_{1},\ldots,\phi_{n} \in A$. Then, $\operatorname{ev}^{-1}(A^{\perp}) = \ker \phi_{1} \cap \ldots \cap \ker \phi_{n}$ which is open in $L$. Now, we check that $\operatorname{ev}$ is injective. Let $v \in L$ be a nonzero vector. Choose a decomposition of $L = U \oplus F$ where $U$ is open and $F$ is finite dimensional containing $v$ (this can be done because $L$ is separated and linearly compact). Let $\phi$ be a linear functional such that $\phi\lvert_{U} = 0$ and $\phi(v) \neq 0$. Since $U$ is open and $F$ discrete such $\phi$ exists and it is continuous. Now, we check that $\operatorname{ev}$ is surjective. Let $\mathscr{U} = \{U_{\lambda}\}_{\lambda\in \Lambda}$ be a system of neighborhoods consisting of open vector subspaces. Let $\psi\colon L^{*} \to k$ be a continuous linear functional. The dual map of $\pi_{\lambda}\colon L \to L/U_{\lambda}$ yields an injection $\pi_{\lambda}^{*}\colon(L/U)^{*} \hookrightarrow L^{*}$ for every $U_{\lambda} \in \mathscr{U}$. Since $L$ is linearly compact, the vector space $L/U_{\lambda}$ is finite dimensional. Thus, there exists a unique $\overline{x_{\lambda}}\in L$ such that $\psi \circ \pi_{\lambda}^{*} = \operatorname{ev}_{\overline{x_{\lambda}}}$ where $\operatorname{ev}\colon L/U_{\lambda} \to (L/U_{\lambda})^{**}$. In addition, observe that if $\mu\leq \lambda $ there is an induced injection $\pi_{\mu}^{\lambda*}\colon(L/U_{\mu})^{*} \hookrightarrow (L/U_{\lambda})^{*}$ such that the following diagram
	\[
	\begin{tikzcd}
		L^{*} \arrow[rr, "\psi"] & & k \\
		(L/U_{\lambda})^{*} \arrow[u, hook, "\pi_{\lambda}^{*}"] \arrow[urr, swap, "\operatorname{ev}_{\overline{x_{\lambda}}}"] & & \\
		(L/U_{\mu})^{*} \arrow[u, hook, "\pi_{\mu}^{\lambda*}"] \arrow[uurr, swap, bend right, "\operatorname{ev}_{\overline{x_{\mu}}}"] \arrow[uu, bend left = 70, hook, "\pi_{\mu}^{*}"]& & 
	\end{tikzcd}
	\]
	commutes. Observe that uniqueness of $\overline{x_{\lambda}} \in L/U_{\lambda}$ implies $\overline{x_\mu} = \pi_{\mu}^{\lambda}(\overline{x_{\lambda}})$ for all $\mu \leq \lambda$. Indeed, for all $\phi_{\mu}\in(L/U_{\mu})^{*}$ the equality
	\begin{align*}
	\phi_{\mu}(\overline{x_{\mu}}) &= \psi(\pi_{\mu}^{*}(\phi_{\mu})) \\
	&= \psi(\pi_{\lambda}^{*} \circ \pi_{\mu}^{\lambda*}(\phi_{\mu})) \\
	&= \psi(\pi_{\lambda}^{*}(\phi_{\mu} \circ \pi_{\mu}^{\lambda})) \\
	&= (\phi_{\mu} \circ \pi_{\mu}^{\lambda})(\overline{x_{\lambda}}) \\
	&= \phi_{\mu}(\pi_{\mu}^{\lambda}(\overline{x_{\lambda}}))
	\end{align*}
	holds. Therefore, $\overline{x_\mu} = \pi_{\mu}^{\lambda}(\overline{x_{\lambda}})$ for all $\mu \leq \lambda$. Then, $(x_{\lambda})_{\lambda\in \Lambda}$ is contained in the completion $\widehat{V}$ as described in \cref{def:completion}. Since $V \to \widehat{V}$ is an isomorphism, there exists $x\in L$ such that $\pi_{\lambda}(x) = \overline{x_{\lambda}}$ for all $\lambda\in \Lambda$. We claim that $\psi = \operatorname{ev}_{x}$. Let $\phi \in L^{*}$. Then, $\ker\phi$ is open and there exists $\lambda\in \Lambda$ such that $U_{\lambda} \subseteq \ker\phi$. Hence, we can factor $\phi$ as follows
	\[
	\begin{tikzcd}
		L \arrow[r, "\phi"] \arrow[d, swap, "\pi_{\lambda}"] & k \\
		L/U_{\lambda} \arrow[ur, swap, "\phi_{\lambda}"] &
	\end{tikzcd}
	\]
	Now, since $L/U$ is discrete it follows that $\phi_{\lambda}$ is continuous. Moreover, $\pi_{\lambda}^{*}(\phi_{\lambda}) = \phi$. Hence
	\[
		\psi(\phi) = \psi(\pi_{\lambda}^{*}(\phi_{\lambda})) = \phi_{\lambda}(\overline{x_{\lambda}}) = \phi_{\lambda}(\pi_{\lambda}(x) = \phi(x).
	\]
	This implies surjectivity of $\operatorname{ev}\colon L \to L^{**}$. Finally, we prove that $\operatorname{ev}$ is open. Let $U$ be any open vector subspace in $L$, thus $L = U \oplus F$ for some finite dimensional $F$. We claim that $\operatorname{ev}(U) = (F^{*})^{\perp}$. First, the inclusion $\operatorname{ev}(U) \subseteq (F^{*})^{\perp}$ is immediate. Let $\psi \in (F^{*})^{\perp}$. Let $x \in L$ such that $\operatorname{ev}_{x} = \psi$. Write $x = u + f$ where $u \in U$ and $f \in F$. Hence, $\operatorname{ev}_{x} = \operatorname{ev}_{u} + \operatorname{ev}_{f}$. Since $\operatorname{ev}$ is injective, it follows that there exists some $\phi\in F^{*}$ such that $\phi(f) \neq 0$ if $f$ is nonzero. Therefore, $f = 0$ and $\psi \in \operatorname{ev}(U)$. This concludes the proof.
\end{proof}
\begin{remark}\label{rem:completeness-is-necessary-for-sel-duality}
	Observe that completeness cannot be dropped in the definition of a Tate space while preserving duality. Indeed, if $V$ is linearly compact but not complete its dual is discrete by \cref{lemm:duality-d-lattice-c-lattice} and by \cref{rem:dual-of-discrete-is-complete} its double dual is complete, hence $V \to V^{**}$ cannot be an isomorphism. In fact, during the proof of the duality theorem we checked that $V^{**}$ is the completion of $V$. 
\end{remark}
\begin{remark}\label{rem:lef-vs-beilinson}
	We now discuss definitions of linearly compact spaces as given in \cite{Lefschetz-Alg-Top} and \cite{Chiral}. In \cite{Lefschetz-Alg-Top} it is proven that a linearly compact vector space must be complete while our definition does not imply it necessarily. However, when $V$ is a complete space both definitions coincide. Indeed, Lefschetz proves that every linearly compact space is the dual of a discrete space which coincides with our definition of a complete linearly compact vector space by \cref{thm:self-duality}. Therefore, his definition of a \textbf{locally linearly compact vector space} (that is, a linearly topologized vector space admitting an open linearly compact vector subspace) coincides with our notion of Tate space.
\end{remark}
\subsection{Morphisms}
\begin{definition}\label{def:linearly-compact-and-discrete-morphisms}
	A morphism $f\colon V\to W$ of Tate spaces is said to be \textbf{linearly compact} if the closure of $f(V)$ is linearly compact in $W$. Dually, it is \textbf{discrete} if $\ker f$ is open in $V$. 
\end{definition}
First, we check the duality natural property for morphisms of Tate spaces.
\begin{proposition}\label{prop:duality-discrete-compact-maps}
	A morphism $f\colon V\to W$ of Tate spaces is linearly compact if and only if $f^{*}$ is discrete.
\end{proposition}
\begin{proof}
	Suppose $f^{*}$ is linearly compact, then $\ker f^{*} = f(V)^{\perp}$. However, if $\phi\in W^{*}$ and $\phi(f(V)) = 0$ then $\phi(\overline{f(V)}) = 0$ by continuity of $\phi$. Therefore, $\ker f^{*} = \overline{f(V)}^{\perp}$ which is open because $\overline{f(V)}$ is linearly compact. Now, suppose $f^{*}$ is discrete. Thus, $\ker f^{*}$ contains a basic open subspace $A^{\perp}$ such that $A$ is linearly compact in $W$. Therefore, $f(V)\subseteq A$ then $\overline{f(V)} \subseteq A$ and by item (c) in \cref{thm:linear_compactness_properties} $\overline{f(V)}$ is linearly compact.
\end{proof}
Discrete and linearly compact operators form a 2-sided ideal in $\Hom$; that is
\begin{proposition}\label{prop:compact-and-discrete-2-sided-ideal}
	If $f$ is a linearly compact morphism (respectively discrete) then its composition with an arbitrary morphism of Tate spaces is also linearly compact (respectively discrete). 
\end{proposition}
\begin{proof}
	Let $A \xrightarrow{f} B \xrightarrow{g} C \xrightarrow{h} D$ be morphisms of Tate spaces such that $g$ is linearly compact. Then, $\overline{g\circ f(A)} \subseteq \overline{g(B)}$ which is linearly compact, thus $\overline{g\circ f(A)}$ is linearly compact as well. On the other hand, note that $h(\overline{g(B)}) \subseteq \overline{h\circ g(B)}$; therefore $\overline{h(\overline{g(B)})} = \overline{h\circ g(B)}$. However, $\overline{g(B)}$ is linearly compact and by item (b) of \cref{thm:linear_compactness_properties} $\overline{h(\overline{g(B)})}$ is linearly compact. In addition, the statement for discrete operators follows from \cref{prop:duality-discrete-compact-maps}.
\end{proof}
\begin{remark}\label{rem:discrete-composition-compact}
If $f$ is a compact operator and $g$ is a discrete operator, then $gf$ is of \textbf{finite-rank}; that is, $\dim gf(V) < \infty$.  
\end{remark}
\begin{proof}
	We have $\overline{f(V)} \prec \ker g$, therefore, $\overline{f(V)}/(\overline{f(V)} \cap \ker g)$ is finite dimensional. We have a surjection
	\[
		\frac{\overline{f(V)}}{\overline{f(V)}\cap \ker g} \to gf(V)
	\]
	which implies that $gf$ is of finite-rank.
\end{proof}
\begin{definition}\label{def:2-sided-ideals-in-hom}
	Let $V$ and $W$ be Tate spaces. \linebreak We denote $\Hom_{+}(V,W)$ to be the set of linearly compact morphisms and $\Hom_{-}(V,W)$ the set of discrete ones. Let
	\[
		\Hom_{0}(V,W):=\Hom_{+}(V,W) \cap \Hom_{-}(V,W).
	\]
\end{definition}
\begin{proposition}\label{prop:discrete-compact-operators-present-the-whole-space}
	The sets $\Hom_{-}(V,W), \Hom_{+}(V,W)$ and \linebreak $\Hom_{0}(V,W)$ are vector subspaces of $\Hom(V,W)$. Moreover,
	\[
		\Hom_{-}(V,W) + \Hom_{+}(V,W) = \Hom(V,W).
	\]
\end{proposition}
\begin{proof}
	Let $L$ be a c-lattice in $V$ and consider $\pi\colon V \to L$ be a continuous linear projection. Then $\pi$ realized as an element of $\End(V)$ satisfies $\pi \in\End_{+}(V)$ and $1 -\pi\in\End_{-}(V)$. Hence, by \cref{prop:compact-and-discrete-2-sided-ideal} for every $f\in \Hom(V,W)$ $f\circ \pi$ and $f\circ (1 - \pi)$ are linearly compact and discrete respectively. It follows
	\[
		\Hom_{-}(V,W) + \Hom_{+}(V,W) = \Hom(V,W).
	\]
	The other statements are immediate.
\end{proof}
\textcolor{red}{I'll include further theory if necessary.}
